A user-extensible and adaptable parser architecture

### 2008-05-27 doc/logparser-acronyms.tex
Tested Splunk.
Finished the architecture section, unless Carl has comments. Lots of minor changes.

\newacronym{SMTP}{Simple Mail Transfer Protocol}{description={
    SMTP is the protocol which transfers mail from the sender to the
    recipient across the Internet.  A brief introduction to SMTP is
    provided in section~\refwithpage{SMTP background}.
}}

@@

\newacronym{SPF}{Sender Policy Framework}{description={
    SPF is introduced in section~\refwithpage{spf introduction} and
    explained fully in~\cite{openspf, Wikipedia-spf}.
}}

@@

>>>>

\newacronym{SMTP}{Simple Mail Transfer Protocol}{description={
    SMTP is the protocol which transfers mail from the sender to the
    recipient across the Internet.  %XXX A brief introduction to SMTP is
    %provided in \sectionref{SMTP background}.
}}

@@

\newacronym{SPF}{Sender Policy Framework}{description={
    SPF is introduced in \sectionref{spf introduction} and explained fully
    in~\cite{openspf, Wikipedia-spf}.
}}

@@

\newacronym{CLI}{Command Line Interface}{description={
    An computer interface based on typing commands, rather than using a
    mouse.
}}

\newacronym{NLP}{Natural Language Processing}{description={
    Natural Language Processing~\cite{Wikipedia-nlp} attempts to increase
    our understanding of the languages normally used by humans (e.g.\
    English, Spanish, Japanese), with the goal of writing programs which
    can understand human languages.  Potential applications include:
    programs which will automatically summarise a text; programs which
    can be taught facts, will understand questions and formulate answers to
    those questions.  It is in no way connected to Neuro-linguistic
    programming, which is generally found when searching for NLP.
}}

### 2008-05-27 doc/logparser-introduction.tex
Finished the architecture section, unless Carl has comments. Lots of minor changes.

This algorithm requires a database for storing both the rules used when
parsing and the results gleaned from parsing.  The database schema used is
described in section~\refwithpage{database schema}, explaining in detail
the tables used for storing the data gleaned from the log files and the
table that stores the rules.

>>>>

This algorithm requires a database for storing both the rules used when
parsing and the results gleaned from parsing.  The database schema used is
described in \sectionref{database schema}, explaining in detail the tables
used for storing the data gleaned from the log files and the table that
stores the rules.

### 2008-05-27 papers/sgai-2008/paper.tex
Finished the architecture section, unless Carl has comments. Lots of minor changes.

% $Id$
\documentclass[]{svmult}

@@

\usepackage[acronym=true,style=altlist,number=none,toc=true]{glossary}
\makeglossary{}
\makeacronym{}

@@

% Replacement for \ref{}, adds the page number too.
\newcommand{\refwithpage}[1]{%
    \empty{}\ref{#1} [page~\pageref{#1}]%
}
% section references, automatically add \textsection
\newcommand{\sectionref}[1]{%
    \textsection{}\refwithpage{#1}%
}

@@

XXX DUE MONDAY JUNE 2$^{ND}$ 2008\@.

PRESENTATION PAPERS\@: MAX 14 PAGES\@.

POSTER PAPERS\@: MAX 6 PAGES\@.

BLACK AND WHITE ONLY\@.

@@

The solution developed is conceptually simple: provide a few generic functions,
each capable of dealing with an entire \textit{category\/} of log lines (e.g.\
rejections), accompanied by a multitude of precise patterns, each of which
matches all log lines of a specific type, and only that type (e.g.\ rejection
by a specific \RBL{}).  XXX STOP \RBL{} BEING EXPANDED AGAIN\@.

XXX WHERE SHOULD I DISCUSS COMPLICATIONS\@?

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome

@@

Explain adding new rules and actions.

Thoughts from Carl:

    \item context-free rules

    \item partially context-aware actions, though not really

    \item transduction: input log lines $\rightarrow$ database

    \item Closer to NLP than fixed grammar.

XXX BEGIN\@:

@@

Decoupling the parsing rules from the associated actions and framework allows
new rules to be written and tested without requiring modifications to the
parser source code (significantly lowering the barrier to entry for new or
casual users who need to parse new log lines), and greatly simplifies
framework, actions and rules. Decoupling the actions from the
framework simplifies both framework and actions: the framework provides
services to the actions, and doesn't need to deal with the complications which
arise, or the task of reconstructing a mail's journey through Postfix; actions
benefit from having services provided by the framework, freeing them to
concentrate on the task of accurately reconstructing each mail's journey
through Postfix and dealing with the complications arising (XXX REFERENCE).

@@

Actions contain the code required to deal with log lines from a particular
category, e.g.\ rejections.  The actions are parser-specific: each parser
will need to write the actions it requires from scratch, unless it's
extending an existing parser.  In the Postfix parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules.

\showgraph{build/plot-action-distribution}{Distribution of rules
per action}{Distribution of rules per action}

@@

    \item [regex] The \regex{} to match the log line against.  The \regex{}
        will first have several keywords expanded: this simplifies reading
        and writing rules; avoids needless repetition of complex \regex{}
        components; allows the components to be corrected and/or improved
        in one location; and makes each \regex{} largely self-documenting.

        For efficiency the keywords are expanded and every rule's \regex{}
        is compiled before attempting to parse the log file --- otherwise
        each \regex{} would be recompiled each time it was used, resulting
        in a large, data dependent slowdown.  Rule efficiency concerns are
        discussed in \sectionref{rule efficiency}, with the impact of
        compiling and caching \regexes{} covered in \sectionref{Caching
        each regex}.

@@

    \item [action] The action that will be invoked when this rule matches a
        log line; actions are describe in \sectionref{Actions}.

@@

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.\footnote{It may be possible to parse
each rule's regex and determine if any overlap.  The author has not
attempted to do this: given the tremendous power and flexibility of
regex, particularly Perl's (XXX REFERENCE FOR PERL\@; GLOSSARY ENTRY?)
regex, it would be an extremely difficult task.} Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the order in which rules are tried against each line may change between log
files, and the first matching rule wins.  Overlapping rules are frequently
a requirement, allowing a more specific rule to match some log lines and a
more general rule to match the majority, e.g.\ separating \SMTP{} delivery
to specific sites from \SMTP{} delivery to the rest of the world.  The
algorithm provides a facility for ordering overlapping rules: the priority
field in each rule (defaults to zero).  Rules are sorted by priority,
highest first, and then rules with the same priority are sorted by the
number of successful matches when parsing the previous log file.  Negative
priorities may be useful for catchall rules.

@@

@@

@@

% Redefine the command used to produce the glossary title, because the
% default command produces an unnumbered section whereas I want a numbered
% section.
\renewcommand{\glossarytitle}{\section{Glossary}\label{Glossary}}
\printglossary{}
% Redefine the command a second time to produce acronyms instead of a
% glossary.
\renewcommand{\glossarytitle}{\section{Acronyms}\label{Acronyms}}
\printacronym{}
\end{document}

>>>>

% $Id$
\documentclass[draft]{svmult}

@@

\usepackage[style=altlist,number=none,toc=true]{glossary}
\makeglossary{}

@@

% Replacement for \ref{}, adds the page number too.
\newcommand{\refwithpage}[1]{%
    \empty{}\ref{#1} [p.~\pageref{#1}]%
}
% section references, automatically add \textsection
\newcommand{\sectionref}[1]{%
    \textsection{}\refwithpage{#1}%
}

@@

XXX DUE MONDAY JUNE 2$^{ND}$ 2008\@.  PRESENTATION PAPERS\@: MAX 14
PAGES\@.  POSTER PAPERS\@: MAX 6 PAGES\@.  BLACK AND WHITE ONLY\@.

@@

The solution developed is conceptually simple: provide a few generic functions,
each capable of dealing with an entire \textit{category\/} of log lines (e.g.\
rejections), accompanied by a multitude of precise patterns, each of which
matches all log lines of a specific type, and only that type (e.g.\ rejection
by a specific \RBLshort{}).

@@

@@

Decoupling the parsing rules from the associated actions and framework allows
new rules to be written and tested without requiring modifications to the
parser source code (significantly lowering the barrier to entry for new or
casual users who need to parse new log lines), and greatly simplifies
framework, actions and rules. Decoupling the actions from the
framework simplifies both framework and actions: the framework provides
services to the actions, and doesn't need to deal with the complications which
arise, or the task of reconstructing a mail's journey through Postfix; actions
benefit from having services provided by the framework, freeing them to
concentrate on the task of accurately reconstructing each mail's journey
through Postfix and dealing with the complications arising (discussed in
\sectionref{complications}).

@@

Actions contain the code required to deal with log lines from a particular
category, e.g.\ rejections.  The actions are parser-specific: each parser
will need to write the actions it requires from scratch, unless it's
extending an existing parser.  In the Postfix parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules.
The distribution of rules to actions is very skewed, as can be seen in
Fig.~\refwithpage{Distribution of rules per action}.
\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} The distribution is skewed for
the same reason the architecture is so successful in enabling users to
parse new inputs: most new inputs can be parsed by a combination of a new
rule and an existing action.  Unsurprisingly the action with the highest
number of associated rules is \texttt{REJECTION}, the action which handles
rejections; it's followed by \texttt{INFO}, the action responsible for
handling informative inputs: these are log lines that supplement the data
gathered from other log lines, but that don't themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: there are many log lines which are not of interest to the
parser, e.g.\ warning messages about \DNS{} mismatches, or log lines
lacking a way to identify which data structure they should be associated
with.\footnote{Postfix allocates a queueid (explained in the
glossary,~\sectionref{Glossary}) to each mail; most log lines contain the
queueid of the associated mail, but some log lines, particularly with older
versions of Postfix, lack a queueid, preventing their association with a
particular data structure.}  The parser warns about every unparsed input,
to alert the user that they need to alter or extend their rules, hence the
\texttt{IGNORE} action: the input has been parsed, but it is of no benefit
so it is ignored.  The remaining actions have only one or two associated
rules: in some cases this is because there will only ever be one type of
input for that action, e.g.\ all log lines showing that a remote client has
connected are matched by a single rule and handled by the \texttt{CONNECT}
action; in other cases the action is required to address a deficiency in
the log files, or a complication which arises during
parsing.\footnote{\label{complications footnote}The complications which
arose during development of the parser will be documented in detail in the
author's thesis.}

The ability to add special purpose actions to deal with difficulties and
new requirements which are discovered during parser development is one of
the strengths of this architecture: instead of writing one monolithic
function which must be modified to support new behaviour or resolve
challenges, with all the attendant risks of adversely affecting the
existing, working parser, a new, independent action can be written to
satisfy the new requirements.  In some cases the new action will require
the cooperation of other actions, e.g.\ to set or check a flag; this does
introduce the possibility introducing failure by modifying existing
actions, but modifications should be smaller and less frequent, and thus
failures should be less likely and easier to test for and diagnose.  The
parser can be written in an Object
Oriented~\cite{Wikipedia-object-orientation} style, allowing sub-classes to
extend or override actions in addition to adding new actions; because each
action is an independent function, the sub-class need only modify the
action it is interested in, rather than reproducing large chunks of
functionality.

\label{complications}

During development of the parser it became apparent that in addition to the
obvious variety in log lines, there were a large number of complications to
be overcome.\footref{complications footnote}  Some were the result of
deficiencies in Postfix's logging, which were improved or resolved by later
versions of Postfix; others were due to the vagaries of process scheduling,
client behaviour, and administrative actions.  All were successfully
accommodated in the parser: adding new actions was enough for several of
the complications; others required modifications to existing actions to
work around the difficulties; the remainder were resolved by adapting
actions to pass information to subsequent actions (via the framework).

@@

    \item [regex] The \regex{} to match the log line against.  The \regex{}
        will first have several keywords expanded: this simplifies reading
        and writing rules; avoids needless repetition of complex \regex{}
        components; allows the components to be corrected and/or improved
        in one location; and makes each \regex{} largely self-documenting.
        For efficiency the keywords are expanded and every rule's \regex{}
        is compiled before attempting to parse the log file --- otherwise
        each \regex{} would be recompiled each time it was used, resulting
        in a large, data dependent slowdown.  Rule efficiency concerns are
        discussed in \sectionref{rule efficiency}, with the impact of
        compiling and caching \regexes{} covered in \sectionref{Caching
        each regex}.

@@

    \item [action] The action that will be invoked when this rule matches a
        log line; actions are described in \sectionref{Actions}.

@@

Adding new rules is simple: simply define all the required attributes and
append the new rule to wherever the existing rules are stored.  The
software developed for this project includes a utility based on
\SLCT{}~\cite{slct-paper} to aid in producing \regexes{} to parse
previously unparsed inputs.

@@

XXX DO I NEED A SECTION ABOUT OVERLAPPING RULES\@?

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.\footnote{It may be possible to parse each
rule's regex and determine if any overlap.  The author has not attempted to
do this: given the tremendous power and flexibility of regexes,
particularly Perl's regexes~\cite{perlre}, it would be an extremely
difficult task.} Unintentionally overlapping rules lead to inconsistent
parsing and data extraction because the order in which rules are tried
against each line may change between log files, and the first matching rule
wins.  Overlapping rules are frequently a requirement, allowing a more
specific rule to match some log lines and a more general rule to match the
majority, e.g.\ separating \SMTP{} delivery to specific sites from \SMTP{}
delivery to the rest of the world.  The algorithm provides a facility for
ordering overlapping rules: the priority field in each rule (defaults to
zero).  Rules are sorted by priority, highest first, and then rules with
the same priority are sorted by the number of successful matches when
parsing the previous log file.  Negative priorities may be useful for
catchall rules.

@@

\subsection{Architecture characteristics}

\label{Architecture characteristics}

The architecture can be compared to existing approaches to aid in fully
understanding it.

    \item [Context-free rules:]  Rules are context free: they do not take
        into account inputs which have previously been seen, nor do they
        perform any kind of lookahead to future inputs.  In context-free
        grammar terms the parser rules could be described as:

        $\text{\textless{}log-line\textgreater{}} \mapsto \text{rule-1} |
        \text{rule-2} | \text{rule-3} | \dots | \text{rule-n}$

    \item [Partially context-sensitive actions:]  XXX IS THIS THE CORRECT
        TERMINOLOGY\@?  Actions can consult the results of previous actions
        when executing, providing some context sensitivity.  In some cases
        the lack of results from previous actions can be useful
        information, e.g.\ in the Postfix parser when a client disconnects
        if there have been no useful log lines for that client the data is
        discarded, rather than trying to store empty records in the
        database.

    \item [Matching is simple:]  The first matching rule wins: there is no
        backtracking to try alternate rules, no attempt made to pick a
        \textit{best\/} rule; the first rule which matches the line is the
        rule which parses it and determines the action that will be
        invoked.\footnote{Rule authors can affect the order rules will be
        tried in by setting the \texttt{priority} attribute of one or more
        rules.}

    \item [Transduction:]  The parser can be thought of as implementing
        transduction: it takes data in one form (log files) and transforms
        it to another form: a database in the case of the Postfix log
        parser, though it could be normalised files or files which only
        contain discrepancies or other interesting data.

    \item [Closer to \NLP{} than fixed grammar:]  Unlike traditional
        parsers such as those used when compiling a programming language,
        this architecture does not have a fixed grammar specification which
        inputs must adhere to.  The architecture is designed for easy user
        alteration, in fact such alteration is expected.

@@

The architecture's greatest strength is the ease with which new inputs
can be parsed and new actions written.

@@

% Redefine the command used to produce the glossary title, because the
% default command produces an unnumbered section whereas I want a numbered
% section.
\renewcommand{\glossarytitle}{\section{Glossary}\label{Glossary}}
\printglossary{}

\glossary{name={queueid},description={
    Each mail in Postfix's queue is assigned a queueid to uniquely identify
    it.  Queueids are assigned from a limited pool, so although they are
    guaranteed to be unique for the lifetime of the mail, given sufficient
    time they will be reused.
}}

\end{document}
